Causal Bandits: Learning Good Interventions via Causal Inference

### 2016-02-05 causal-bandit/icml-2015/experiments.tex
updates experiments section

We compare the performance of the Algorithm 1, which is specific to the parallel problem, but does not require knowledge of $\boldsymbol{q}$, with that of Algorithm 2 and the Successive Reject algorithm of \cite{}. For each experiment, the results are over 10,000 simulation and error bars show three standard errors.

In figure \ref{fig:simple_vs_m} we fix the number of variables $N$ and the horizon $T$ and compare the performance of the algorithms as $m$ increases. The regret for the Successive Reject algorithm is constant as it depends only on the reward distribution and has no knowledge of the causal structure. For the causal algorithms it increases approximately with $\sqrt{m}$. As $m$ approaches $N$, the gain the causal algorithms obtain from knowledge of the structure is outweighed by fact they do not take into account the observed rewards to focus sampling effort on actions with likely high pay-offs.

\begin{figure}[h]
\caption{Simple regret vs $m(\boldsymbol{q})$ for fixed horizon $T=300$ and number of variables $N = 50$}
\label{fig:simple_vs_m}
\centering
\includegraphics[width=.5\textwidth]{exp_regret_vs_m_N50_T400_s100_20160205_1637.pdf}
\end{figure}

If figures \ref{fig:simple_vs_T_vary_epsilon} and \ref{fig:simple_vs_T} we compare the performance as a function of $T$. \todof{finish this}

@@

\begin{figure}
\caption{Simple regret vs horizon, $T$, with $N = 50$, $m=2$ and $\epsilon = \sqrt{K/4T}$}
\label{fig:simple_vs_T_vary_epsilon}
\centering
\includegraphics[width=.5\textwidth]{exp_regret_vs_T_N50_a4_s10000_20160204_1831}
\end{figure}

@@

\subsection{Highly confounded problem}

\begin{figure}[h]
\centering
\caption{Causal model for the highly confounded problem.}
\label{fig:causalStructure_highly_confounded}
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1cm,
  thick,main node/.style={observed}, hidden/.style={empty}]
\node[main node](1){$V_{2}$};
\node[main node, right=of 1](2){$X_{3}$};
\node[hidden, right=of 2](3){$...$};
\node[main node, right=of 3](4){$X_{N}$};
\node[main node, below right=of 2](5){Y};
\node[main node,above right=of 2](6){$X_1$}; 
 \path[every node/.style={font=\sffamily\small}]
    (1) edge (5)
    	(2) edge (5)
    (4) edge (5)
    (6) edge (1) edge (2) edge (4);
\end{tikzpicture}
\end{figure}

A model under which the value of the confounding variable $X_1$ plays the central role in determining $P(Y)$.  Let:
\eq {
\P{Z = 1} &= q \\
\P{X_k = 1|Z=j} &=  q_j \;\; \forall k \in \set{1...n}\\
\P{Y|x_1 ... x_n} &= \bar{\boldsymbol{x}}
} 
This leads to rewards:
\eq {
\P{Y|do(Z = j)} = q^j(1-q)^{1-j}\\
\P{Y|do(X_i = j)} = \begin{cases}
\frac{1}{n} + \frac{n-1}{n}\left(q q_1 + (1-q)q_0 \right) & \text{ if } j = 1 \\
\frac{n-1}{n}\left(q q_1 + (1-q)q_0 \right) & \text { if } j = 0
\end{cases}
}

>>>>

We compare the performance of the Algorithm 1, which is specific to the parallel problem, but does not require knowledge of $\boldsymbol{q}$, with that of Algorithm 2 and the Successive Reject algorithm of \cite{}. For each experiment, we show the average regret over 10,000 simulations with error bars displaying three standard errors.

In figure \ref{fig:simple_vs_m} we fix the number of variables $N$ and the horizon $T$ and compare the performance of the algorithms as $m$ increases. The regret for the Successive Reject algorithm is constant as it depends only on the reward distribution and has no knowledge of the causal structure. For the causal algorithms it increases approximately with $\sqrt{m}$. As $m$ approaches $N$, the gain the causal algorithms obtain from knowledge of the structure is outweighed by fact they do not take into account the observed rewards to focus sampling effort on actions with high pay-offs.

\begin{figure}[h]
\caption{Simple regret vs $m(\boldsymbol{q})$ for fixed horizon $T=400$ and number of variables $N = 50$}
\label{fig:simple_vs_m}
\centering
\includegraphics[width=.5\textwidth]{exp_regret_vs_m_N50_T400_s1000_20160205_0910.pdf}
\end{figure}

Figure \ref{fig:simple_vs_T_vary_epsilon} demonstrates the performance of the algorithms in the worst case environment for standard bandits, where the gap between the optimal and sub-optimal arms,$\epsilon \sim \sqrt{\frac{N}{8T}}$ , is just too small to be learned. This gap is learn-able by the causal algorithms, for which the worst case depends on the $m \\l N$. In figure \ref{fig:simple_vs_T} we fix $N$ and $\epsilon$ and observe that, for sufficiently large $T$, the regret decays exponentially. The decay constant is larger for the causal algorithms as they have observed a greater effective number of samples for a given $T$.

@@

\begin{figure}
\caption{Simple regret vs horizon, $T$, with $N = 50$, $m=2$ and $\epsilon = \sqrt{\frac{N}{8T}}$}
\label{fig:simple_vs_T_vary_epsilon}
\centering
\includegraphics[width=.5\textwidth]{exp_regret_vs_T_N50_a4_s10000_20160204_1831}
\end{figure}

@@
